Assessing the Global COVID19 Impact on Air Transport with Open Data

### 2021-06-18 paper.Rmd
up until worldwide data, without the regions yet

1.2.1 Gather and prepare data
```{r}
#Reading raw data protocol

source("./R/list_apt_files.R")
year <- "2021"
file_names <- list_apt_files(.year = year)
open_sky <- map_dfr(file_names, read_csv)
fb1 <- open_sky %>% transmute(typecode = as.factor(typecode), origin = as.factor(origin), destination = as.factor(destination), firstseen, lastseen) %>% drop_na()

#write_csv(open_sky, "./data/open_sky.csv")

#Reading prepared data
open_sky <- read_csv("data/open_sky.csv")
```

```{r}
# Take a look
glimpse(open_sky)
str(open_sky)
summary(open_sky)

# First exploration
fb1 <- open_sky %>% transmute(typecode = as.factor(typecode), origin = as.factor(origin), destination = as.factor(destination), firstseen, lastseen)
str(fb1)
summary(fb1)
# More than one million NA's

fb2 <- fb1 %>% drop_na()
str(fb2)
summary(fb2)
# Better...

#Visualizing something...
df1 <- fb2 %>% group_by(DATE = date(firstseen)) %>% summarize(count = n())
df1 %>% ggplot(aes(x = as.factor(year(DATE)))) +
  geom_boxplot(aes(y = count))
# This is the daily traffic of Aprils, distributed as a boxplot, FOR THE WHOLE WORLD! Needs to filter.

>>>>

1.2.1 Gather and prepare data
```{r}
#Reading raw data

source("./R/list_apt_files.R")
year <- "202105"
file_names <- list_apt_files(.year = year)
open_sky <- map_dfr(file_names, read_csv)

#Selecting relevant variables

fb <- open_sky %>% transmute(ADEP = origin, ADES = destination, TYPE = as.factor(typecode), DATE = date(day), CALL = callsign, ACFT_ID = aircraft_uid)

#Easily "dropping NA's# - this can be further sofisticated

fb1 <- fb %>% drop_na()
fb1
```

```{r}
#Joining coutries and regions
# NOTE: the airports.csv file below is the downloaded file from www.ourairports.org.

airports <- read_csv("data-raw/airports.csv") %>% transmute(ICAO = ident, CTRY = iso_country)
#There are missing airports
airports <- airports %>% add_row(ICAO  = c("SPJC", "YBMC", "LSZM", "YSCH", "EPLB", "K3M3", "VV01", "SC28", "CWF2", "EHDB", "74XA", "HE13"), CTRY  = c("PE", "AU", "FR", "AU", "PL", "US", "VN", "US", "CA", "NL", "US", "EG"))

#Joining to ADEP
fb2 <- left_join(fb1, airports, by = c("ADEP" = "ICAO")) %>% mutate(ADEP_CTRY = CTRY, .keep = "unused", .after = ADEP)
#Joining to ADES
fb3 <- left_join(fb2, airports, by = c("ADES" = "ICAO")) %>% mutate(ADES_CTRY = CTRY, .keep = "unused", .after = ADES)

#Check  NA's
colSums(is.na(fb3))
# Very few NA's - it's safe to drop and factor now

world_dataset <- fb3 %>% drop_na() %>% mutate(ADEP_CTRY = as.factor(ADEP_CTRY), ADES_CTRY = as.factor(ADES_CTRY))
world_dataset
```

```{r}
